### 2015-12-11 article/preamble-latex.tex
complete rewrite using standard ropensci article format

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,tocloft,titlesec}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% format headings
\titleformat{\subsection}
{\normalfont\Large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\normalfont\large\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titleformat{\subparagraph}
{\normalfont\normalsize\itshape}{\thesubparagraph}{1em}{}

\titlespacing\paragraph{0pt}{12pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}
\titlespacing{\subparagraph}{0pt}{6pt plus 4pt minus 2pt}{+0pt plus 2pt minus 2pt}

% format toc
\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}