### 2016-05-02 code/rmarkdown/tables.Rmd
+ fix bug with missing species variable being overwritten + assorted figures and tables fixes

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

```

## Tables

```{r, echo=FALSE}
curr.DF <- correlation.model.summary.DF %>%
	select(-contains('raw')) %>%
	mutate(Environmental.corr.P=pvalString(Environmental.corr.P),
		Geographic.corr.P=pvalString(Geographic.corr.P), blank='', Species=gsub('\\ ', '\\n', Species)) %>%
	select(Species:Environmental.corr.P, blank, Geographic.Marginal.R2:Geographic.corr.P) %>%
	format.dataframe(omit=c('Environmental.corr.P', 'blank', 'Geographic.corr.P')) 
emphasize.strong.cells(which(curr.DF$Environmental.corr.P < 0.05, arr.ind=TRUE))
emphasize.strong.cells(which(curr.DF$Geographic.corr.P < 0.05, arr.ind=TRUE))
curr.txt=knitr::kable(
	curr.DF,
	digits=2,
	caption='Correlation between surrogate variables and genetic variation. The $R^2$ columns show the proportion of genetic variation explained by surrogates. Specifically, marginal $R^2$ represents the proportion of variation explained by the fixed effects alone. The conditional $R^2$ represents the proportion of variation explained by the fixed and random effects. All $\\chi^2$ tests were based on one degree of freedom. Significant $P$ values ($P$ < 0.05) indicate that surrogate variables correlate with genetic variation.',
	col.names=c('\\textbf{Species}', '\\textbf{Marginal $R^2$}', '\\textbf{Conditional $R^2$}', '\\textbf{$\\chi^2$}', '\\textbf{\\textit{P}}', '', '\\textbf{Marginal $R^2$}', '\\textbf{Conditional $R^2$}', '\\textbf{$\\chi^2$}', '\\textbf{\\textit{P}}'),
	align=rep('c', ncol(curr.DF)),
	format='latex'
) %>% as.character()

curr.txt <- gsub('\\hline\nSpecies',
'
\\hline
& \\multicolumn{4}{c}{Environmental} & & \\multicolumn{4}{c}{Geographic} \\\\ \\cmidrule{2-5} \\cmidrule{7-10}
Species', curr.txt, fixed=TRUE)

class(curr.txt) <- 'knitr_kable'
attr(curr.txt, 'format') <- 'latex'
curr.txt
```