### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ add diagrams to supplementary figures

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble.tex, si-preamble.tex]
---
# Supporting Information
All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
# load .rda file
checkpoint::checkpoint('2016-08-03', R.version='3.3.1', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
grid.sub.FPLY <- spFortify(grid.sub.PLY)

# define ggplot2 color wheel function
gg_color_hue <- function(n) {
	hues <- seq(15, 375, length=n+1)
	hcl(h=hues, l=65, c=100)[1:n]
}
```

# Figures

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap=paste0('Species distributions. Squares represent planning units. For a given species, planning units that it was found in are shown in bright blue.')}
## plot map of species distributions
# fortify data
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
) %>% mutate(presence=factor(c('absent', 'present')[presence+1]))

# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', face='italic'),
		axis.title=element_blank()
	) +
		scale_fill_manual(name='Occupancy', values=c('absent'='#FC8D59', 'present'='#91CF60')) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	) +
	facet_wrap(~ species, ncol=4)
```

```{r, fig.width=15, fig.height=3, fig.cap='Climatic variation. Each panel depicts variation based on a principle component (PC) for a different species. Squares represent planning units. The color of each planning unit denotes the average PC value of pixels inside it. For a given species, the planning units with more similar colors have more similar climate regimes.', echo=FALSE}
plot(1)
```

```{r, echo=FALSE}
# prepare data
pop.probs.DF <- llply(
	seq_along(unique(spp.samples.DF$species)),
	function(i) {
		curr.pts <- filter(
			spp.sample.PTS@data,
			species==unique(spp.samples.DF$species)[i]
		) %>% mutate(
			Individual=seq_along(species)
		) 	
		curr.probs <- structurer:::sample.membership(spp.StructureCollection.LST[[i]]) %>%
			as.data.frame() %>%
			mutate(
				species=unique(spp.samples.DF$species)[i],
				Individual=seq_along(species)) %>%
			gather(Population,Probability,-species,-Individual) %>%
			mutate(Population=as.numeric(gsub('V', '', Population, fixed=TRUE))) %>%
			left_join(curr.pts, by=c('species','Individual')) %>%
			mutate(species=gsub('_', ' ', species, fixed=TRUE)) %>%
			arrange(species,Individual,Population)
		return(curr.probs)
	}
) %>% do.call(what=rbind)
# create legend colors
legend.CHR <- `names<-`(
	gg_color_hue(n_distinct(pop.probs.DF$Population)),
	as.character(unique(pop.probs.DF$Population))
)
# create function
plot.spp.pop.probs <- function(i) {
	# init spp.samples
	prop <- 0.85
	curr.spp.DF <- filter(pop.probs.DF, species==unique(pop.probs.DF$species)[i])
	# create base map grob
	curr.GROB <- ggplot(data=curr.spp.DF) +
		geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
			fill='grey85', color='grey75') +
		geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
			group='group'), fill='grey70', color='grey10') +
		guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
		theme_classic() +
		theme(
			axis.title=element_blank(), axis.text=element_blank(),
			axis.line=element_blank(), axis.ticks=element_blank(),
			axis.ticks.length = unit(0, "cm"),
			legend.position = "none", panel.margin = unit(0, "lines"),
			plot.margin = unit(c(0, 0, 0, 0), "lines"),
			strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white', face='italic', size=20),
			legend.text= element_text(size=20)
		) + 
		coord_cartesian(
			xlim=buffered.range(grid.FPLY$long, 0),
			ylim=buffered.range(grid.FPLY$lat, 0)
		) +
		scale_fill_manual(
			name='',
			values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
		)
	# create header grob
	curr.GROBH <- ggplot() +
			geom_text(aes(x=0,y=0,label=species),
				data=curr.spp.DF[1,,drop=FALSE],
				color='white', fontface='italic', size=4) +
		theme_classic() +
		theme(
			axis.title=element_blank(), axis.text=element_blank(),
			axis.line=element_blank(), axis.ticks=element_blank(),
			axis.ticks.length = unit(0, "cm"),
			legend.position = "none", panel.margin = unit(0, "lines"),
			plot.margin = unit(c(0, 0, 0, 0), "lines"),
			panel.background=element_rect(fill='grey20')
		) +
		coord_cartesian(xlim=c(-1,1), ylim=c(-1,1))
	# create bbox object
	bbox <- c(ggplot_build(curr.GROB)$panel$ranges[[1]]$x.range,
		ggplot_build(curr.GROB)$panel$ranges[[1]]$y.range)
		names(bbox) <- c('ll.lon', 'ur.lon', 'll.lat', 'ur.lat')
	# render grob
	grid.newpage()
	pushViewport(viewport(layout=grid.layout(2,1, heights=unit(c(1-prop, prop), "npc"))))
	print(curr.GROBH, vp=viewport(layout.pos.row=1, layout.pos.col=1))
	print(curr.GROB, vp=viewport(layout.pos.row=2, layout.pos.col=1))
	# add in pies
	for (j in unique(curr.spp.DF$cell)) {
		# init
		curr.cell.DF <- filter(curr.spp.DF, cell==j) %>% 
			group_by(cell, Population) %>%
			summarize(
				grid.longitude=mean(grid.longitude), 
				grid.latitude=mean(grid.latitude),
				Probability=mean(Probability)
			) %>% data.frame()
		curr.lon <- (curr.cell.DF$grid.longitude[1]-bbox['ll.lon'])/abs(bbox['ur.lon']-bbox['ll.lon'])
		curr.lat <- ((curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat']))*prop
		# create pie chart for individual
		curr.pie <- ggplot(curr.cell.DF, aes(x="", y=Probability, fill=factor(Population))) +
			geom_bar(width=0.75, color='grey20', stat='identity') +
			coord_polar(theta='y') +
			theme(
				line = element_blank(), rect = element_blank(), 
				text = element_blank(), axis.ticks.length = unit(0, "cm"),
				legend.position = "none", panel.margin = unit(0, "lines"),
				plot.margin = unit(c(0, 0, 0, 0), "lines"), complete = TRUE
			) +
			scale_fill_manual(
				name='',
				values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
			)
		# render pie
		print(curr.pie, vp=viewport(x=curr.lon, y=curr.lat, width=0.075, height=0.075))
	}
}
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_',' ',unique(spp.samples.DF$species)[1], fixed=TRUE),'} populations. Squares denote planning units. Pie charts denote quadrats where individuals were sampled, and colors denote the average probability that individuals in the quadrat belong to different populations on average.')}
plot.spp.pop.probs(1)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[2],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (2 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(2)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[3],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (3 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(3)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[4],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (4 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(4)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[5],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (5 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(5)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[6],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (6 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(6)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[7],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (7 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(7)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[8],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (8 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(8)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[9],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (9 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(9)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[10],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (10 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(10)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[11],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(11)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[12],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (12 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(12)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[13],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (13 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(13)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[14],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (14 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(14)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[15],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (15 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(15)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[16],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (16 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(16)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[17],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (17 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(17)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[18],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (18 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(18)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[19],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (19 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(19)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[20],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (20 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(20)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[21],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (21 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(21)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[22],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (22 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(22)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[23],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (23 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(23)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[24],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (24 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(24)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[25],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (25 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(25)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.cap=paste0('Distribution of \\textit{',gsub('_', ' ',unique(spp.samples.DF$species)[26],fixed=TRUE),'} populations. See Figure S3 caption for conventions.')}
if (26 <= general.params.LST[[MODE]]$n.spp) plot.spp.pop.probs(26)
```

```{r, echo=FALSE, include=FALSE}
plot.spp.nmds <- function(i) {
	## prepare data
	# get vector of adaptive and neutral column names
	adapt.col <- grep(
		paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
		names(grid.FPLY), value=TRUE, fixed=TRUE
	)
	neutral.col <- grep(
		paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
		names(grid.FPLY), value=TRUE, fixed=TRUE
	)
	# skip species if not used in analysis
	curr.FPLY <- grid.FPLY[,
		c(adapt.col, neutral.col,
		'long','lat','group'),drop=FALSE]
	curr.FPLY <- curr.FPLY[which(!is.na(curr.FPLY[[1]])),]
	curr.FPLY <- curr.FPLY %>% gather(d, value, contains('adaptive'), contains('neutral')) %>% mutate(
		d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE),
		d=gsub('_', ' ', d, fixed=TRUE)
	)
	for (j in seq_len(length(adapt.col)+length(neutral.col)))
		curr.FPLY[['d']] <- gsub(as.character(j), paste0('(', j, ')'), curr.FPLY[['d']], fixed=TRUE)
	curr.FPLY <- curr.FPLY %>% mutate(d=factor(gsub(' d(', ' (', d, fixed=TRUE)))
	# make plot
	curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey85', color='grey70') +
			geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat', 
				group='group', fill='value'), color='grey10') +
			theme_classic() +
			theme(
				axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
				panel.border=element_rect(color='black', fill=NA, size=1),
				strip.background = element_rect(fill='grey20'), strip.text = element_text(color='white', size=20)
			) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0),
				ylim=buffered.range(grid.FPLY$lat, 0)
			) +
			scale_fill_gradientn(name='',
				colors=brewer.pal(9, 'RdYlBu'),
				guide=guide_colorbar(ticks=element_line(color='black'),
					border=element_line(color='black'), barheight=unit(3.0, 'cm'))
			) + facet_wrap(~d)
	})
	do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
	invisible()
}

calc.fig.height <- function(i) {
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	n.plots <- 0 
	# count adaptive plot
	if (adapt.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
	# count neutral plot
	if (neutral.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
	# exports
	if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
	return(1.85)
}

calc.fig.width <- function(i) {
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	n.plots <- 0 
	# count adaptive plot
	if (adapt.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
	# count neutral plot
	if (neutral.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
	# exports
	if (n.plots >= 3) return(10)
	return(4 * n.plots)
}

is.speciesPlottable <- function(i) {
	if (i > general.params.LST[[MODE]]$n.spp)
		return(FALSE)
	return(!is.na(unique(spp.samples.DF$species)[i]))
}
```

```{r, echo=FALSE, fig.width=calc.fig.width(1), fig.height=calc.fig.height(1), fig.width=calc.fig.width(1), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[1]),'}. Each panel shows a different dimension of the variation as indicated by its title. Squares represent planning units. The color of each planning unit panel corresponds to the average ordinated values associated with individuals sampled inside it. No indiviudals of this species were found inside the gray planning units.')}
if (is.speciesPlottable(1)) plot.spp.nmds(1)
 ```

```{r, echo=FALSE, fig.width=calc.fig.width(2), fig.height=calc.fig.height(2), fig.width=calc.fig.width(2), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[2]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(2)) plot.spp.nmds(2)
```

```{r, echo=FALSE, fig.width=calc.fig.width(3), fig.height=calc.fig.height(3), fig.width=calc.fig.width(3), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[3]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(3)) plot.spp.nmds(3)
```

```{r, echo=FALSE, fig.width=calc.fig.width(4), fig.height=calc.fig.height(4), fig.width=calc.fig.width(4), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[4]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(4)) plot.spp.nmds(4)
```

```{r, echo=FALSE, fig.width=calc.fig.width(5), fig.height=calc.fig.height(5), fig.width=calc.fig.width(5), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[5]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(5)) plot.spp.nmds(5)
```

```{r, echo=FALSE, fig.width=calc.fig.width(6), fig.height=calc.fig.height(6), fig.width=calc.fig.width(6), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[6]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(6)) plot.spp.nmds(6)
```

```{r, echo=FALSE, fig.width=calc.fig.width(7), fig.height=calc.fig.height(7), fig.width=calc.fig.width(7), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[7]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(7)) plot.spp.nmds(7)
```

```{r, echo=FALSE, fig.width=calc.fig.width(8), fig.height=calc.fig.height(8), fig.width=calc.fig.width(8), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[8]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(8)) plot.spp.nmds(8)
```

```{r, echo=FALSE, fig.width=calc.fig.width(9), fig.height=calc.fig.height(9), fig.width=calc.fig.width(9), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[9]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(9)) plot.spp.nmds(9)
```

```{r, echo=FALSE, fig.width=calc.fig.width(10), fig.height=calc.fig.height(10), fig.width=calc.fig.width(10), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[10]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(10)) plot.spp.nmds(10)
```

```{r, echo=FALSE, fig.width=calc.fig.width(11), fig.height=calc.fig.height(11), fig.width=calc.fig.width(11), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[11]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(11)) plot.spp.nmds(11)
```

```{r, echo=FALSE, fig.width=calc.fig.width(12), fig.height=calc.fig.height(12), fig.width=calc.fig.width(12), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[12]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(12)) plot.spp.nmds(12)
```

```{r, echo=FALSE, fig.width=calc.fig.width(13), fig.height=calc.fig.height(13), fig.width=calc.fig.width(13), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[13]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(13)) plot.spp.nmds(13)
```

```{r, echo=FALSE, fig.width=calc.fig.width(14), fig.height=calc.fig.height(14), fig.width=calc.fig.width(14), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[14]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(14)) plot.spp.nmds(14)
```

```{r, echo=FALSE, fig.width=calc.fig.width(15), fig.height=calc.fig.height(15), fig.width=calc.fig.width(15), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[15]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(15)) plot.spp.nmds(15)
```

```{r, echo=FALSE, fig.width=calc.fig.width(16), fig.height=calc.fig.height(16), fig.width=calc.fig.width(16), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[16]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(16)) plot.spp.nmds(16)
```

```{r, echo=FALSE, fig.width=calc.fig.width(17), fig.height=calc.fig.height(17), fig.width=calc.fig.width(17), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[17]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(17)) plot.spp.nmds(17)
```

```{r, echo=FALSE, fig.width=calc.fig.width(18), fig.height=calc.fig.height(18), fig.width=calc.fig.width(18), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[18]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(18)) plot.spp.nmds(18)
```

```{r, echo=FALSE, fig.width=calc.fig.width(19), fig.height=calc.fig.height(19), fig.width=calc.fig.width(19), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[19]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(19)) plot.spp.nmds(19)
```

```{r, echo=FALSE, fig.width=calc.fig.width(20), fig.height=calc.fig.height(20), fig.width=calc.fig.width(20), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[20]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(20)) plot.spp.nmds(20)
```

```{r, echo=FALSE, fig.width=calc.fig.width(21), fig.height=calc.fig.height(21), fig.width=calc.fig.width(21), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[21]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(21)) plot.spp.nmds(21)
```

```{r, echo=FALSE, fig.width=calc.fig.width(22), fig.height=calc.fig.height(22), fig.width=calc.fig.width(22), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[22]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(22)) plot.spp.nmds(22)
```

```{r, echo=FALSE, fig.width=calc.fig.width(23), fig.height=calc.fig.height(23), fig.width=calc.fig.width(23), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[23]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(23)) plot.spp.nmds(23)
```

```{r, echo=FALSE, fig.width=calc.fig.width(24), fig.height=calc.fig.height(24), fig.width=calc.fig.width(24), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[24]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(24)) plot.spp.nmds(24)
```

```{r, echo=FALSE, fig.width=calc.fig.width(25), fig.height=calc.fig.height(25), fig.width=calc.fig.width(25), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[25]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(25)) plot.spp.nmds(25)
```

```{r, echo=FALSE, fig.width=calc.fig.width(26), fig.height=calc.fig.height(26), fig.width=calc.fig.width(26), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[26]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(26)) plot.spp.nmds(26)
```

```{r, echo=FALSE, fig.width=calc.fig.width(27), fig.height=calc.fig.height(27), fig.width=calc.fig.width(27), fig.cap=paste0('Distribution of adaptive and neutral genetic variation in \\textit{',gsub('\\_', ' ', unique(spp.samples.DF$species)[27]),'}. See Figure S29 caption for conventions.')}
if (is.speciesPlottable(27)) plot.spp.nmds(27)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=(2.5*4), fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environmental variation it also secures. Points represent values associated with randomly generated prioritizations. Blue lines indicate average trends computed using a generalised linear model with a logit link. Each panel shows data for a different species.'}
# format data
curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
curr.DF2$surrogate.held <- 0
curr.DF2$genetic.held <- 0.9

# make plot
suppressWarnings(
	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
		theme_classic() +
		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
			panel.margin.x=unit(1, 'lines')) +
		xlab('Environmental variation secured (%)') +
		ylab('Adaptive genetic variation secured (%)') +
		geom_hex() +
		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
			se=FALSE, fill='blue', alpha=0.8)) +
		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
		facet_wrap(~ Species, ncol=4)
)
```

```{r, include=FALSE}
calc.fig.height3 <- function(x) {
	return((((x-1) %/% 4)+1)*1.5)
}

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=calc.fig.height3(n_distinct(geo.surrogacy.DF$Species)), fig.cap='The relationship between the proportion of neutral genetic variation secured in a prioritizaiton and the proportion of geographic variation it also secures. Conventions are detailed in Figure S56.'}
# format data
curr.DF <- geo.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2 <- geo.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
curr.DF2$surrogate.held <- 0
curr.DF2$genetic.held <- 0.9

# make plot
suppressWarnings(
	ggplot(aes(x=surrogate.held, y=genetic.held), data=curr.DF) +
		theme_classic() +
		theme(strip.background=element_rect(fill='grey20'), strip.text=element_text(color='white', face='italic', size=12),
			panel.margin.x=unit(1, 'lines')) +
		xlab('Geographic variation secured (%)') +
		ylab('Neutral genetic variation secured (%)') +
		geom_hex() +
		suppressWarnings(stat_smooth(method='glm', method.args=list(family='binomial'), fullrange=TRUE,
			se=FALSE, fill='blue', alpha=0.8)) +
		geom_text(aes(label=R2_repr), data=curr.DF2, parse=TRUE, hjust=0, size=5, color='red') +
		facet_wrap(~ Species, ncol=4)
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies. Squares represent planning units, and their colors indicate their selection frequency. Each column of panels shows solutions generated using different targets. Each row of panels shows solutions generated under different scenarios. Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species. The selection frequencies for the multi-species prioritizations are each based on a single optimal pioritization for all species and so are binary.'}
# prepare data for plotting
grid.sub.FPLY <- grid.sub.PLY
for (i in seq_along(single.spp.prioritisations[[1]]))
	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(multi.spp.prioritisations))
	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
for (i in seq_along(multi.spp.prioritisations.with.cost))
	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
grid.sub.FPLY <- spFortify(grid.sub.FPLY)
grid.sub.FPLY <- grid.sub.FPLY %>% gather(
		Combination,
		`Selection.Frequency`,
		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
	) %>% mutate(
		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
	) %>% mutate(
		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
		Context=factor(
			gsub(' (', '\n(', Context, fixed=TRUE),
			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
	)

# make maps
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
		group='group', fill='Selection.Frequency'),
		alpha=0.8, color=NA) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white', size=8),
		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
		legend.title=element_text(vjust=-5),
		legend.margin=unit(c(0.1), 'cm'),
		axis.title=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.sub.FPLY$long, 0),
		ylim=buffered.range(grid.sub.FPLY$lat, 0)
	) +
	facet_grid(Context ~ Prioritization) +
	scale_fill_gradient(name='Selection Frequency (%)',
		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
			ticks=element_line(color='black'), border=element_line(color='black'))
	)
```

```{r, include=FALSE}
render_gvis <- function(x) {
  file = tempfile(fileext='.png')
  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(x)), file)
  grid::grid.raster(png::readPNG(file))
}
```

```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Overall data processing workflow. The flowcharts show the methods used to generate the environmental and geographic surrogate spaces, and the adaptive and neutral genetic spaces. Squares indicate data sets. Ellipses indicate analyses. Lines show relationships between data and analysis, and their captions describe how data was processed.'}
render_gvis(grViz("
digraph main {
  graph [layout = dot]

  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  meirmans_data [label = <Plant data<BR></BR>(Meirmans <I>et al. </I> 2011)>]
  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
  sample_location_data [label = 'Sample location data']
  grid_data [label = 'Grid cell data']
  sample_aflp_data [label = 'Sample AFLP data']

  geo_as [label = 'Geographic attribute spaces'] 
  env_as [label = 'Environmental attribute spaces']
  neutral_as [label = 'Neutral genetic attribute spaces']
  adaptive_as [label = 'Adaptive genetic attribute spaces']

  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
  calc_geo_as [label = 'Processing\n(Figure SXXX)'] 
  calc_env_as [label = 'Processing\n(Figure SXXX)']
  calc_genetic_as [label = 'Processing\n(Figure SXXX)']

  edge [color = grey]
  meirmans_data -> {sample_aflp_data grid_data sample_location_data}
  bioclimatic_data -> calc_env_as
  sample_location_data -> {calc_geo_as calc_env_as calc_genetic_as}
  grid_data -> {calc_geo_as calc_env_as calc_genetic_as}
  sample_aflp_data -> calc_genetic_as
  calc_genetic_as -> {neutral_as, adaptive_as}
  calc_geo_as -> geo_as
  calc_env_as -> env_as
  calc_env_as -> calc_env_as [label = ' each species\nseparately']
  calc_geo_as -> calc_geo_as [label = ' each species\nseparately']
  calc_genetic_as -> calc_genetic_as [label = ' each species\nseparately']

}
"))
```

```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
render_gvis(grViz("
digraph geo_as {
  graph [layout = dot]

  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  # input
  sample_location_data [label = 'Sample location data']
  grid_data [label = 'Grid cell data']
  species_name [label = 'Species name']
  {rank = same; sample_location_data; grid_data; species_name}

  # intermediate
  spp_samples [label = 'Locations of samples for species']
  spp_grids [label = 'Grid cells occupied by species']
  spp_proj_data [label = 'Projected grid cells']
  spp_centriod_data [label = 'Centroids of grid cells']

  # output
  geographic_space [label = 'Geographic attribute space']

  edge [color = grey]

  species_name -> spp_samples 
  sample_location_data -> spp_samples [label = ' subset data']
  spp_samples -> spp_grids [label = ' overlay data']
  grid_data -> spp_grids 
  spp_grids -> spp_proj_data [label = ' reproject data to\nequi-distant CRS']
  spp_proj_data -> spp_centriod_data [label = ' compute centroids']
  spp_centriod_data -> geographic_space [label =' extract coordinates and use them\nto represent demand points and\nplanning unit coordinates']

}
"))
```

```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
render_gvis(grViz("
digraph env_as {
  graph [layout = dot]

  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  # input
  sample_location_data [label = 'Sample location data']
  grid_data [label = 'Grid cell data']
  species_name [label = 'Species name']
  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
  {rank = same; sample_location_data; grid_data; species_name; bioclimatic_data}

  # intermediate
  spp_samples [label = 'Locations of samples for species']
  spp_grids [label = 'Grid cells occupied by species']
  spp_proj_data [label = 'Projected grid cells']

  bioclimatic_proj_data [label = 'Projected bioclimatic data']
  clipped_bioclimatic_data [label = 'Bioclimatic data clipped to species distribution']
  pca_scores_data [label = 'Main axes of bioclimatic variation\nacross species distribution']
  grid_pca_data [label = 'Environmental conditions inside each grid cell']

  # output
  environmental_space [label = 'Environmental attribute space']

  # analyses
  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]  
  pca [label = 'Principal components analysis']

  edge [color = grey]
  species_name -> spp_samples 
  sample_location_data -> spp_samples [label = ' subset data']
  grid_data -> spp_grids
  bioclimatic_data -> bioclimatic_proj_data [label = ' reproject data to\nequal-area CRS']
  spp_samples -> spp_grids [label = ' overlay data']
  spp_grids -> spp_proj_data [label = ' reproject data to\nequal-area CRS']
  bioclimatic_proj_data -> clipped_bioclimatic_data [label = ' clip data']
  spp_proj_data -> clipped_bioclimatic_data
  clipped_bioclimatic_data -> pca
  pca -> pca_scores_data [label = ' extract scores for\nfirst two components']
  spp_proj_data -> grid_pca_data
  pca_scores_data -> grid_pca_data [label = ' calculate average scores\nfor each grid cell']
  grid_pca_data -> environmental_space [label =' extract scores and use them\nto represent demand points and\nplanning unit coordinates']

}
"))
```

```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S59 caption.'}
render_gvis(grViz("
digraph genetic_as {
  graph [layout = dot]

  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  # input
  sample_location_data [label = 'Sample location data']
  sample_aflp_data [label = 'Sample AFLP data']
  grid_data [label = 'Grid cell data']
  species_name [label = 'Species name']
  {rank = same; sample_location_data; grid_data; species_name; sample_aflp_data}

  # intermediate
  spp_aflp_samples [label = 'Species AFLP data']
  spp_location_samples [label = 'Species location data']
  spp_grids [label = 'Grid cells occupied by species']

  population_data [label = 'Samples grouped into populations']
  subset_data [label = 'Samples with reliable population designations']
  rare_alleles [label = 'AFLP data for\nrare alleles']
  remainder_alleles [label = 'AFLP data for\n standard alleles']
  bayescan_adaptive_alleles [label = ' Outlier loci']
  pcadapt_adaptive_alleles [label = ' Outlier loci']
  bayescan_neutral_alleles [label = ' Standard loci']
  pcadapt_neutral_alleles [label = ' Standard loci']
  adaptive_alleles [label = ' AFLP data\nfor adaptive loci']
  neutral_alleles [label = ' AFLP data\nfor neutral loci']
  adaptive_ordinations [label = ' Ordinations showing main\ngradients of variation\nin adaptive loci']
  neutral_ordinations [label = '  Ordinations showing main\ngradients of variation\nin neutral loci']
  grid_adaptive_ordinations [label = ' Typical adaptive genetic characteristics of individuals in grid cells']
  grid_neutral_ordinations [label = ' Typical neutral genetic characteristics of individuals in grid cells']

  # output
  adaptive_genetic_space [label = 'Adaptive genetic space']
  neutral_genetic_space  [label = 'Neutral genetic space']

  # analyses
  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
  bayescan_analysis [label = 'BayeScan analysis']
  pcadapt_analysis [label = 'PCadapt analysis']
  structure_analysis [label = 'Structure analysis']
  clumpp_analysis [label = 'ClumPP analysis']
  neutral_nmds_analysis [label = 'NMDS analyses']
  adaptive_nmds_analysis [label = 'NMDS analyses']
  combine_results [label = 'Combine results from all analyses.\nLoci classified as adaptive in both analyses are treated as adaptive,\notherwise treated as neutral']

  edge [color = grey]
  sample_location_data -> spp_location_samples [label = ' subset data']
  species_name -> spp_location_samples 
  sample_aflp_data -> spp_aflp_samples [label = ' subset data']
  species_name -> spp_aflp_samples
  spp_location_samples -> spp_grids
  grid_data -> spp_grids [label = ' overlay data']
  spp_aflp_samples -> structure_analysis
  structure_analysis -> clumpp_analysis [label = ' post-process\nresults']
  clumpp_analysis -> population_data
  population_data -> subset_data [label = ' omit samples within\nuncertain memberships']
  subset_data -> rare_alleles [label = ' extract rare alleles where\nglobal frequency of minor\nalleles < 10 %']
  subset_data -> remainder_alleles
  remainder_alleles -> bayescan_analysis [label = 'use population\ngroupings']
  remainder_alleles -> pcadapt_analysis [label = 'ignore population\ngroupings']
  bayescan_analysis -> bayescan_adaptive_alleles [label = ' extract']
  pcadapt_analysis -> pcadapt_adaptive_alleles  [label = ' extract']
  bayescan_analysis -> bayescan_neutral_alleles [label = ' extract']
  pcadapt_analysis -> pcadapt_neutral_alleles [label = ' extract']
  bayescan_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
  pcadapt_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
  bayescan_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
  pcadapt_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
  rare_alleles -> combine_results [label = ' treat loci\nas neutral']
  combine_results -> adaptive_alleles
  combine_results -> neutral_alleles 
  neutral_alleles  -> neutral_nmds_analysis
  adaptive_alleles -> adaptive_nmds_analysis
  adaptive_nmds_analysis -> adaptive_ordinations [label = ' extract']
  neutral_nmds_analysis  -> neutral_ordinations [label = ' extract']
  spp_grids  -> grid_adaptive_ordinations
  adaptive_ordinations -> grid_adaptive_ordinations [label = ' calculate average values\nin each grid cell']
  spp_grids  -> grid_neutral_ordinations
  neutral_ordinations -> grid_neutral_ordinations [label = ' calculate average values\nin each grid cell']
  grid_adaptive_ordinations -> adaptive_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
  grid_neutral_ordinations -> neutral_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']

}
"))
```

# Tables

```{r, echo=FALSE}
write.table(loci_classification.DF, 'article/Table_S1.csv', sep=',', row.names=FALSE)
```

```{r, echo=FALSE, results='asis'}
adaptive.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='adaptive') %>% mutate(Stress=as.character(round(Stress,3))) %>% mutate(Stress=replace(Stress, Stress=='0', '< 0.001'))
neutral.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='neutral') %>% mutate(Stress=as.character(round(Stress,3))) %>% mutate(Stress=replace(Stress, Stress=='0', '< 0.001'))
names(adaptive.spp.nmds.summary.DF)[2:5] <- paste0('adaptive.', names(adaptive.spp.nmds.summary.DF)[2:5])
names(neutral.spp.nmds.summary.DF)[2:5] <- paste0('neutral.', names(neutral.spp.nmds.summary.DF)[2:5])
spp.nmds.summary.DF2 <- full_join(adaptive.spp.nmds.summary.DF, neutral.spp.nmds.summary.DF, by='Species')
spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)],'}')
spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)],'}')
spp.nmds.summary.DF2 <- spp.nmds.summary.DF2 %>%
	mutate(blank='') %>%
	select(Species, adaptive.k:adaptive.Stress, neutral.k:neutral.Stress)
curr.file <- tempfile(fileext='.tex')
x <- spp.nmds.summary.DF2 %>%
	latex(
		file=curr.file,
		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 5),
		cgroup=c('', 'Adaptive', 'Neutral'), n.cgroup=c(1, 2, 2),
		colheads=c('\\bfseries{Species}', 'K', 'Stress', 'K', 'Stress'), here=TRUE,
		caption='Summary of non-metric multi-dimensional scaling (NMDS) analyses on genetic variation for each species. The analyses were used to identify the main graondients of variation in binary adaptive and neutral loci data. Bold stress values indicate NMDS analyses that have converged.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

```{r, echo=FALSE, results='asis'}
curr.file <- tempfile(fileext='.tex')
x <- correlation.model.summary.DF %>%
	select(-contains('raw')) %>%
	mutate(
		Environmental.corr.P=pvalString(Environmental.corr.P),
		Geographic.corr.P=pvalString(Geographic.corr.P),
		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
		blank=rep('', length(Species)),
		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
	latex(file=curr.file, 
		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
		caption='Summary of maximum likelihood population-effects (MLPE) models. Briefly, for each species associated with potentially adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group). Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group). Data shows the results from comparing each model to its corresponding null model using $\\mathbf{\\chi^2}$ tests (1 degree of freedom), and the $\\mathbf{R^2}$ ($\\mathbf{MR^2}$) and conditional $\\mathbf{R^2}$ ($\\mathbf{CR^2}$) for each model.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
	gsub(pattern='\\toprule', replacement='', fixed=TRUE) %>%
	gsub(pattern='\\midrule', replacement='\\midrule', fixed=TRUE) %>%
	gsub(pattern='\\bottomrule', replacement='\\midrule', fixed=TRUE) %>%
	cat()
```

```{r, echo=FALSE, results='asis'}
curr.file <- tempfile(fileext='.tex')
x <- surrogate.spp.DF %>%
	mutate(
		Environmental.z = paste(round(Environmental.estimate,2), '$\\pm$', round(Environmental.std.error,2)),
		Geographic.z = paste(round(Geographic.estimate,2), '$\\pm$', round(Geographic.std.error,2)),
		Species = paste0('\\textit{', gsub('_', ' ', Species, fixed=TRUE), '}')) %>%
	mutate(
		Environmental.z=replace(Environmental.z, grepl('NA', Environmental.z, fixed=TRUE),''),
		Geographic.z=replace(Geographic.z, grepl('NA', Geographic.z, fixed=TRUE),'')
	) %>%
	select(Species, Environmental.z, Environmental.p.value, Environmental.r2CU,
		Geographic.z, Geographic.p.value, Geographic.r2CU) %>%
	latex(file=curr.file, 
		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 3, 3), col.just=rep('c', 7),
		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
		colheads=c('\\bfseries{Species}', '$\\textit{Z}$', '\\textit{P}', '$R^2$', '$\\textit{Z}$', '\\textit{P}', '$R^2$'),
		caption='Summary of post-hoc analyses on generalized linear models. Briefly, two generalized linear models were fit to correlate the proportion of adaptive genetic variation and environmental variation secured, and the proportion of neutral genetic variation and geographic variation secured in a prioritization. These models also included a predictor variable indicating the species for which the prioritisations were generated, and an interaction term. Table shows the results from one-sided _post-hoc_ analyses that were conducted to determine if the slopes for each species were positive (using the alternative hypothesis that slopes were less than or equal to zero; tests used a single degree of freedom; results reported in the \\textit{Z} and \\textit{P} columns). The table also shows Cragg and Uhler\'s pseudo-R$^2$ values that were calculated to show the proportion of variation explained by the slopes for each species ($R^2$ column).'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{2-5} \\cline{7-', fixed=TRUE) %>% 
	gsub(pattern='ulticolumn{1}{c}{$R^2$}\\tabularnewline', replacement='ulticolumn{1}{c}{$R^2$}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

```{r, echo=FALSE, results='asis'}
## make table showing pca resuls
curr.file <- tempfile(fileext='.tex')
x <- spp.pca.DF %>%
	mutate(species=gsub('_', ' ', species, fixed=TRUE), percent_variation_explained=percent_variation_explained*100) %>%
	latex(
		file=curr.file,
		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 2),
		colheads=c('\\bfseries{Species}', '\\bfseries{Variation explained (\\%)}'), here=TRUE,
		caption=paste0('Summary of principal component analyses fit to the climatic variation in the species\' ranges. Percentages correspond to the percent of variation described by the first ',as.character(as.english(surrogate.params.LST[[MODE]]$number.components)),' principle components.')
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```