### 2016-11-10 code/rmarkdown/si-preamble.tex
+ assorted changes

% load packages
\usepackage{amsmath,amsfonts,float,makecell,natbib,lineno,booktabs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

% insert S before figure and table caption
\pagenumbering{gobble}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{1}