### 2017-10-18 code/rmarkdown/slides.Rmd

---
fontsize: 11pt
documentclass: article
output:
  rmarkdown::pdf_document
---

# Slides

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo"), dpi = 500)

```

```{r,include=FALSE}
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
  countriesHigh$ADMIN %in% c(
    'Italy', 'Switzerland', 'France', 'Austria',
    'Germany', 'Slovenia', 'Croatia', 'Hungary',
    'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
  )
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
grid.sub.FPLY <- spFortify(grid.sub.PLY)

# define ggplot2 color wheel function
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length=n+1)
  hcl(h=hues, l=65, c=100)[1:n]
}
```

```{r, echo=FALSE, fig.align='center', fig.width=6, fig.height=1.5, out.width='3.42in', height='2.5in', fig.cap="Map of the study area showing (a) the pattern of species richness, and (b) the distribution of acquisition cost, both plotted on a quantile-based color ramp. Squares denote planning units."}

# calculate species richness
grid.PLY$Species_richness <- grid.PLY@data %>%
  select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()

panel.DF <- grid.sub.FPLY %>%
  summarize(lat=min(lat)+(diff(range(lat))*0.96),
    long=min(long)+(diff(range(long))*0.04))

# plot species richness
p1 <- ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat,
    group=group, fill=Species_richness), color=NA) +
  scale_fill_gradientn(name='Richness',
    colors=brewer.pal(5, 'YlGn'),
    guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
  ) +
  theme_classic() +
  theme(
    axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
    panel.border=element_rect(color='black', fill=NA, size=1),
    panel.background=element_rect(color='black', fill='black'),
    plot.background=element_rect(color='black', fill='black'),
    legend.key=element_rect(color='black', fill='black'),
    legend.background=element_rect(color='black', fill='black'),
    legend.title=element_text(color='white'),
    legend.text=element_text(color='white')
  ) +
  coord_cartesian(
    xlim=buffered.range(grid.FPLY$long, 0),
    ylim=buffered.range(grid.FPLY$lat, 0)
  )

qn01 <- scales::rescale(quantile(grid.PLY@data$pop.density, seq(0,1,length.out=9), na.rm=TRUE))
p2 <- ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat,
    group=group, fill=pop.density), color=NA) +
  scale_fill_gradientn(name='Cost',
    colors=brewer.pal(5,'YlOrRd'),
    values=qn01,
    guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
  ) +
  theme_classic() +
  theme(
    axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
    panel.border=element_rect(color='black', fill=NA, size=1),
    panel.background=element_rect(color='black', fill='black'),
    plot.background=element_rect(color='black', fill='black'),
    legend.key=element_rect(color='black', fill='black'),
    legend.background=element_rect(color='black', fill='black'),
    legend.title=element_text(color='white'),
    legend.text=element_text(color='white')
  ) +
  coord_cartesian(
    xlim=buffered.range(grid.FPLY$long, 0),
    ylim=buffered.range(grid.FPLY$lat, 0)
  )

grid.arrange(p1,p2,nrow=1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap=paste0('The proportion of adaptive (a; ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)),' species) and neutral (b; ',sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' species) genetic variation represented in a suite of randomly generated prioritizations for different species as a function of the surrogate variation they also secured (lines). Panels (c) and (d) show the distribution of $R^2$ values for the ',sum(!is.na(correlation.model.summary.DF$Environmental.corr.P)) + sum(!is.na(correlation.model.summary.DF$Geographic.corr.P)),' models shown in panels (a) and (b).')}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
  xlab('Environmental variation\nsecured (%)') +
  ylab('Adaptive genetic\nvariation secured (%)') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  coord_cartesian(xlim=c(0, 1)) +
  scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
  scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
for (i in seq_along(env.surrogacy.LST))
  p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
    data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
    fullrange=TRUE, se=FALSE, colour='white')
# p1 <- p1 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(a)')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
  xlab('Geographic variation\nsecured (%)') +
  ylab('Neutral genetic\nvariation secured (%)') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  coord_cartesian(xlim=c(0, 1)) +
  scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
  scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
for (i in seq_along(geo.surrogacy.LST))
  p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
    data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
    fullrange=TRUE, se=FALSE, colour='white')
# p2 <- p2 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(b)')

p1.hist <- ggplot(aes(r2CU),
    data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
  geom_histogram(fill='white', bins=12) +
  ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  # geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(c)') +
  scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
  scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})

p2.hist <- ggplot(aes(r2CU),
    data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
  geom_histogram(fill='white', bins=12) +
  ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black')) +
  # geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(d)') +
  scale_y_continuous(labels=function(x) {formatC(x, digits=2, format='f')}) +
  scale_x_continuous(limits=c(0, 1), labels=function(x) {formatC(x, digits=2, format='f')})

# render figure
grid.arrange(p1,p2,p1.hist,p2.hist,nrow=2)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=3.5, out.width='4in', out.height='3.5in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios. Box plots show the median, and the 25th and 75th percentiles. Whiskers show the data at 1.5 times the inter-quantile range. Points show data outside the whiskers. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>%
  group_by(Prioritisation, Context, Metric, Species) %>%
  summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
  data.frame() %>%
  filter(Context == 'Multi-species\n(equal costs)')

# make plot
suppressWarnings({
  ggplot(aes(x=Prioritisation,y=genetic.held),
    data=scenario.SDF) +
    geom_boxplot(position=position_dodge(0.9), color='white', fill='black') +
    scale_fill_manual(name='Prioritization',
      values=c('grey80','grey50','grey20')) +
    geom_abline(slope=0, color='red', linetype='dashed',
      intercept=raptr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
    ylab('Genetic variation secured (%)') +
    xlab('') +
    scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
    theme_classic() +
    theme(
         axis.line.x=element_line(color='white'),
         axis.line.y=element_line(color='white'),
         axis.title=element_text(color='white'),
         axis.ticks=element_line(color='white'),
         panel.background=element_rect(color='black',fill='black'),
         plot.background=element_rect(color='black',fill='black'),
         strip.background = element_rect(fill='grey20'),
         strip.text = element_text(color='white', size=9),
         panel.border = element_rect(color='white', fill=NA),
         axis.text.x=element_text(color='white',angle=45, hjust=1, vjust=1,
                                  size=8),
         axis.text.y=element_text(color='white',size=8)) +
    facet_grid(Context ~ Metric)
})
```

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap=paste0('Species distributions. Squares represent planning units. For a given species, planning units that it was found in are shown in bright blue.')}
## plot map of species distributions
# fortify data
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
    z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
    names(z)[4] <- 'presence'
    z$species <- gsub('\\_', ' ', x)
    return(z)
  }
) %>% mutate(presence=factor(c('absent', 'present')[presence+1]))

# plot species data
ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat,
    group=group, fill=presence), alpha=0.8, color='grey10') +
  theme_classic() +
  theme(
    axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
    strip.background = element_rect(fill='grey20'),
    strip.text = element_text(color='white', face='italic'),
    axis.title=element_blank(),
    panel.background=element_rect(color='black',fill='black'),
    plot.background=element_rect(color='black',fill='black'),
    legend.text=element_text(color='white'),
    legend.title=element_text(color='white'),
    legend.key=element_rect(fill='black'),
    legend.background=element_rect(fill='black')
  ) +
    scale_fill_manual(name='Occupancy', values=c('absent'='#FC8D59', 'present'='#91CF60')) +
  coord_cartesian(
    xlim=buffered.range(grid.FPLY$long, 0),
    ylim=buffered.range(grid.FPLY$lat, 0)
  ) +
  facet_wrap(~ species, ncol=4)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(6.5*0.8), out.width='3.42in', height='3.9in', fig.cap='Planning unit selection frequencies. Squares represent planning units, and their colors indicate their selection frequency. Each column of panels shows solutions generated using different targets. Each row of panels shows solutions generated under different scenarios. Note that the selection frequencies for the single-species prioritizations are based on their frequency among prioritizations generated for each species. The selection frequencies for the multi-species prioritizations are each based on a single optimal prioritization for all species and so are binary.'}
# prepare data for plotting
grid.sub.FPLY <- grid.sub.PLY
for (i in seq_along(single.spp.prioritisations[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
  grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(multi.spp.prioritisations))
  grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
for (i in seq_along(multi.spp.prioritisations.with.cost))
  grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
grid.sub.FPLY <- spFortify(grid.sub.FPLY)
grid.sub.FPLY <- grid.sub.FPLY %>% gather(
    Combination,
    `Selection.Frequency`,
    `Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
  ) %>% mutate(
    Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
    Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
  ) %>% mutate(
    Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
    Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
    Context=factor(
      gsub(' (', '\n(', Context, fixed=TRUE),
      levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
        "Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
  )

# make maps
ggplot() +
  geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
    fill='grey85', color='grey70', size=0.001) +
  geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat',
    group='group', fill='Selection.Frequency'),
    alpha=0.8, color=NA) +
  theme_classic() +
  theme(axis.ticks=element_blank(), axis.text=element_blank(),
    plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
    strip.background = element_rect(fill='grey20'),
    strip.text = element_text(color='white', size=8),
    legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
    legend.title=element_text(vjust=-5),
    legend.margin=unit(c(0.1), 'cm'),
    axis.title=element_blank(),
    legend.title=element_text(color='white'),
    legend.text=element_text(color='white'),
    legend.key=element_rect(fill='black'),
    legend.background=element_rect(fill='black'),
    panel.background=element_rect(fill='black'),
    plot.background=element_rect(fill='black')
  ) +
  coord_cartesian(
    xlim=buffered.range(grid.sub.FPLY$long, 0),
    ylim=buffered.range(grid.sub.FPLY$lat, 0)
  ) +
  facet_grid(Context ~ Prioritization) +
  scale_fill_gradient(name='Selection Frequency (%)',
    guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
      ticks=element_line(color='black'), border=element_line(color='black'))
  )
```

```{r, include = FALSE}
plot.spp.spaces <- function(i) {
  ## prepare data
  # get vector of adaptive and neutral column names
  adapt.col <- grep(
    paste0(unique(spp.samples.DF$species)[i], '_adaptive_d'),
    names(grid.FPLY), value=TRUE, fixed=TRUE
  )
  neutral.col <- grep(
    paste0(unique(spp.samples.DF$species)[i], '_neutral_d'),
    names(grid.FPLY), value=TRUE, fixed=TRUE
  )
  env.col <- grep(
    paste0(unique(spp.samples.DF$species)[i], '_env_d'),
    names(grid.FPLY), value=TRUE, fixed=TRUE
  )
  # skip species if not used in analysis
  curr.FPLY <- grid.FPLY %>%
    select(one_of(adapt.col, neutral.col, env.col), group, long, lat) %>%
    mutate_each(funs(unname), one_of(adapt.col, neutral.col, env.col)) %>%
    filter(!is.na(.[[1]])) %>%
    gather(d, value, contains('adaptive'), contains('neutral'), contains('env')) %>%
    mutate(d=as.character(d)) %>%
    mutate(d=gsub(paste0(unique(spp.samples.DF$species)[i], '_'), '', d, fixed=TRUE)) %>%
    mutate(d=gsub('_', ' ', d, fixed=TRUE)) %>%
    mutate(d=gsub('env', 'environmental', d, fixed=TRUE))
  for (j in seq_len(max(length(adapt.col),length(neutral.col),length(env.col))))
    curr.FPLY$d <- gsub(paste0('d', as.character(j)), paste0('(', j, ')'), curr.FPLY$d, fixed=TRUE)
  # make plot
  curr.GROBS <- llply(unique(curr.FPLY$d), function(x) {
    ggplot() +
      geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
        fill='grey85', color='grey70', size=0.001) +
      geom_polygon(data=filter(curr.FPLY, d==x), aes_string(x='long', y='lat',
        group='group', fill='value'), color='grey10') +
      theme_classic() +
      theme(
        axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
        panel.border=element_rect(color='black', fill=NA, size=1),
        strip.background = element_rect(fill='grey20'),
        strip.text = element_text(color='white', size=20),
        legend.title=element_text(color='white'),
        legend.text=element_text(color='white'),
        legend.key=element_rect(fill='black'),
        legend.background=element_rect(fill='black'),
        panel.background=element_rect(fill='black'),
        plot.background=element_rect(fill='black')
      ) +
      coord_cartesian(
        xlim=buffered.range(grid.FPLY$long, 0),
        ylim=buffered.range(grid.FPLY$lat, 0)
      ) +
      scale_fill_gradientn(name='',
        colors=brewer.pal(9, 'RdYlBu'),
        guide=guide_colorbar(ticks=element_line(color='black'),
          border=element_line(color='black'), barheight=unit(3.0, 'cm')),
        labels=function(x) {formatC(x, format='f', digits=2)}
      ) + facet_wrap(~d)
  })
  do.call(grid.arrange, append(curr.GROBS, list(ncol=3)))
  invisible()
}

calc.fig.height <- function(i) {
  # init
  adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
  neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), fixed=TRUE) %>%
    sum()

  n.plots <- 0
  # count adaptive plot
  if (adapt.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
  # count neutral plot
  if (neutral.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
  # count env plot
  n.plots <- n.plots + env.col
  # exports
  if (n.plots == 0) return(1)
  if (n.plots >= 3) return((((n.plots-1) %/% 3)+1) * 2)
  return(1.85)
}

calc.fig.width <- function(i) {
  # init
  adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
  neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
  env.col <- grepl(paste0(unique(spp.samples.DF$species)[i], '_env_d'), names(grid.FPLY), fixed=TRUE) %>%
    sum()
  n.plots <- 0
  # count adaptive plot
  if (adapt.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
  # count neutral plot
  if (neutral.col %in% names(grid.FPLY))
    n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
  # count env plot
  n.plots <- n.plots + env.col
  # exports
  if (n.plots == 0) return(1)
  if (n.plots >= 3) return(10)
  return(4 * n.plots)
}
is.speciesPlottable <- function(i) {
  if (i > general.params.LST[[MODE]]$n.spp)
    return(FALSE)
  return(!is.na(unique(spp.samples.DF$species)[i]))
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='Qvalues from PCadapt for \textit{Arabis alpina}'}
# prepare data
alpha_value <- pcadapt.params.LST[[MODE]]$alpha.level
outlier_data <- data.frame(qvalue = spp.pcadapt.LST[[2]]$qvalues$qvalues) %>%
                mutate(id = factor(seq_len(nrow(.)))) %>%
                mutate(type = case_when(.$qvalue < alpha_value ~ 'adaptive',
                                        .$qvalue >= alpha_value ~ 'neutral'))

# make plot
ggplot(data = outlier_data, aes(x = id, y = qvalue, fill = type)) +
  geom_bar(stat='identity', color = 'black') +
  geom_hline(yintercept = alpha_value, color = "red", linetype = "dashed") +
  xlab('Loci') +
  ylab('Q-value') +
  scale_fill_manual(values = c('adaptive' = 'red', 'neutral' = 'white')) +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        legend.text=element_text(color='white'),
        legend.title=element_text(color='white'),
        legend.background=element_rect(color='black', fill = 'black'),
        legend.key=element_rect(color='black', fill = 'black'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='Qvalues from BayeScan for \textit{Arabis alpina}'}
# prepare data
outlier_data <- spp.BayeScan.LST[[2]]$bayescan@results@summary %>%
                mutate(id = factor(seq_len(nrow(.))))
# make plot
ggplot(data = outlier_data, aes(x = id, y = mean_qval, fill = type)) +
  geom_bar(stat = 'identity', color = 'black') +
  xlab('Loci') +
  ylab('Q-value') +
  theme_classic() +
  scale_fill_manual(values = c('adaptive' = 'red', 'neutral' = 'white')) +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        legend.text=element_text(color='white'),
        legend.title=element_text(color='white'),
        legend.background=element_rect(color='black', fill = 'black'),
        legend.key=element_rect(color='black', fill = 'black'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='NMDS for neutral alleles for \textit{Arabis alpina}'}
# prepare data
nmds <- as.data.frame(spp.nmds.LST[[2]]$neutral$points)
# make plot
ggplot(data = nmds, aes(x = MDS1, y = MDS2)) +
  geom_point(color = 'white') +
  xlab('NMDS 1') +
  ylab('NMDS 2') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='NMDS for adaptive alleles for \textit{Arabis alpina}'}
# prepare data
nmds <- as.data.frame(spp.nmds.LST[[2]]$adaptive$points)
# make plot
ggplot(data = nmds, aes(x = MDS1, y = MDS2)) +
  geom_point(color = 'white') +
  xlab('NMDS 1') +
  ylab('NMDS 2') +
  theme_classic() +
  theme(axis.line.x=element_line(color='white'),
        axis.line.y=element_line(color='white'),
        axis.title=element_text(color='white'),
        axis.text=element_text(color='white'),
        axis.ticks=element_line(color='white'),
        panel.background=element_rect(color='black',fill='black'),
        plot.background=element_rect(color='black',fill='black'))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4.5, fig.height=4.5, fig.cap='Population memberships for \textit{Arabis alpina}'}
# prepare data
pop.probs.DF <- llply(
  seq_along(unique(spp.samples.DF$species)),
  function(i) {
    curr.pts <- filter(
      spp.sample.PTS@data,
      species==unique(spp.samples.DF$species)[i]
    ) %>% mutate(
      Individual=seq_along(species)
    )
    curr.probs <- structurer:::sample.membership(spp.StructureCollection.LST[[i]]) %>%
      as.data.frame() %>%
      mutate(
        species=unique(spp.samples.DF$species)[i],
        Individual=seq_along(species)) %>%
      gather(Population,Probability,-species,-Individual) %>%
      mutate(Population=as.numeric(gsub('V', '', Population, fixed=TRUE))) %>%
      left_join(curr.pts, by=c('species','Individual')) %>%
      mutate(species=gsub('_', ' ', species, fixed=TRUE)) %>%
      arrange(species,Individual,Population)
    return(curr.probs)
  }
) %>% do.call(what=rbind)
# create legend colors
legend.CHR <- `names<-`(
  gg_color_hue(n_distinct(pop.probs.DF$Population)),
  as.character(unique(pop.probs.DF$Population))
)
plot.spp.pop.probs <- function(i) {
  # init spp.samples
  prop <- 0.85
  curr.spp.DF <- filter(pop.probs.DF, species==unique(pop.probs.DF$species)[i])
  # create base map grob
  curr.GROB <- ggplot(data=curr.spp.DF) +
    geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
      fill='grey85', color='grey75') +
    geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat',
      group='group'), fill='grey70', color='grey10') +
    guides(fill=guide_legend(title=''), color=guide_legend(title='')) +
    theme_classic() +
    theme(
      axis.title=element_blank(), axis.text=element_blank(),
      axis.line=element_blank(), axis.ticks=element_blank(),
      plot.background=element_rect(fill='black'),
      axis.ticks.length = unit(0, "cm"),
      legend.position = "none", panel.margin = unit(0, "lines"),
      plot.margin = unit(c(0, 0, 0, 0), "lines"),
      strip.background = element_rect(fill='grey20'),
      strip.text = element_text(color='white', face='italic', size=20),
      legend.text= element_text(size=20)
    ) +
    coord_cartesian(
      xlim=buffered.range(grid.FPLY$long, 0),
      ylim=buffered.range(grid.FPLY$lat, 0)
    ) +
    scale_fill_manual(
      name='',
      values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
    )
  # create header grob
  curr.GROBH <- ggplot() +
      geom_text(aes(x=0,y=0,label=species),
        data=curr.spp.DF[1,,drop=FALSE],
        color='white', fontface='italic', size=4) +
    theme_classic() +
    theme(
      axis.title=element_blank(), axis.text=element_blank(),
      axis.line=element_blank(), axis.ticks=element_blank(),
      axis.ticks.length = unit(0, "cm"),
      legend.position = "none", panel.margin = unit(0, "lines"),
      plot.margin = unit(c(0, 0, 0, 0), "lines"),
      panel.background=element_rect(fill='grey20')
    ) +
    coord_cartesian(xlim=c(-1,1), ylim=c(-1,1))
  # create bbox object
  bbox <- c(ggplot_build(curr.GROB)$panel$ranges[[1]]$x.range,
    ggplot_build(curr.GROB)$panel$ranges[[1]]$y.range)
    names(bbox) <- c('ll.lon', 'ur.lon', 'll.lat', 'ur.lat')
  # render grob
  grid.newpage()
  pushViewport(viewport(layout=grid.layout(2,1, heights=unit(c(1-prop, prop), "npc"))))
  print(curr.GROBH, vp=viewport(layout.pos.row=1, layout.pos.col=1))
  print(curr.GROB, vp=viewport(layout.pos.row=2, layout.pos.col=1))
  # add in pies
  for (j in unique(curr.spp.DF$cell)) {
    # init
    curr.cell.DF <- filter(curr.spp.DF, cell==j) %>%
      group_by(cell, Population) %>%
      summarize(
        grid.longitude=mean(grid.longitude),
        grid.latitude=mean(grid.latitude),
        Probability=mean(Probability)
      ) %>% data.frame()
    curr.lon <- (curr.cell.DF$grid.longitude[1]-bbox['ll.lon'])/abs(bbox['ur.lon']-bbox['ll.lon'])
    curr.lat <- ((curr.cell.DF$grid.latitude[1]-bbox['ll.lat'])/abs(bbox['ur.lat']-bbox['ll.lat']))*prop
    # create pie chart for individual
    curr.pie <- ggplot(curr.cell.DF, aes(x="", y=Probability, fill=factor(Population))) +
      geom_bar(width=0.75, color='grey20', stat='identity') +
      coord_polar(theta='y') +
      theme(
        line = element_blank(), rect = element_blank(),
        text = element_blank(), axis.ticks.length = unit(0, "cm"),
        legend.position = "none", panel.margin = unit(0, "lines"),
        plot.margin = unit(c(0, 0, 0, 0), "lines"), complete = TRUE
      ) +
      scale_fill_manual(
        name='',
        values=legend.CHR[seq_len(n_distinct(curr.spp.DF$Population))]
      )
    # render pie
    print(curr.pie, vp=viewport(x=curr.lon, y=curr.lat, width=0.075, height=0.075))
  }
}
plot.spp.pop.probs(2)
```